Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth

### 2020-11-14 analysis/VOCC/2-biomass-depth-only-model.Rmd
recheck for SST effect; plot covariate effects, residuals, and spatial random fields

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    # species: "Redstripe Rockfish"
    # species: "Shortbelly Rockfish"
    region: "All synoptic surveys"
    knots: 500
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    #covs: "-ssid-only" # covariate string used to name model rds 
    covs: "-tv-depth-only"
    #covs: "-trawled-ssid"
    update_model: FALSE
    update_predictions: FALSE
    update_model_check: FALSE
---

@@

covars <- readRDS("data/event-covariates.rds")
data <- dplyr::left_join(biomass, covars)

@@

if (region == "All synoptic surveys") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
}

@@

   try({ imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
    ))

    imm_biomass <- sdmTMB::sdmTMB(
      data = data,
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      # ar1_fields = TRUE, # changed to FALSE July 2020
      # reml = TRUE,
      include_spatial = TRUE,
      nlminb_loops = 2, 
      newton_steps = 2,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass,
      file = paste0("data/", spp, "/mod-imm-biomass-", 
        spp, covs, "-", ssid_string, ".rds"))
   })

  } else {
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

    total_biomass <- sdmTMB::sdmTMB(
      data = data,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      # ar1_fields = TRUE, # all off in July 2020
      include_spatial = TRUE,
      nlminb_loops = 2, 
      newton_steps = 2,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(total_biomass, file = paste0(
      "data/", spp,
      "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
    ))
  }
  tictoc::toc()
}
```

@@

  try({
    imm_biomass <- readRDS(paste0(
      "data/", spp,
      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))

@@

    imm_predictions <- predict(imm_biomass, newdata = nd)
    saveRDS(imm_predictions, file = paste0(
      "data/", spp,
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass.rds"
    ))
  })
} else {
  total_biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

@@

if(params$update_model_check) {
try({
  biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

try({
  biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

@@

saveRDS(point_predictions, file = paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
))
}
```

```{r}
depth_only_predictions <- readRDS(paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
))

@@

try({
  imm_biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-1n3n4n16-prior-FALSE.rds"
  ))
  # imm_biomass<-sdmTMB:::update_model(imm_biomass)
  # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
})

if (!exists("adult_biomass")) {
  try({
    total_biomass <- readRDS(paste0(
      "data/", spp,
      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))
  })
}

@@

```{r}
png(
  file = paste0(
    "figs/", spp,
    "/depth-", spp, covs, ".png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 6
)
gridExtra::grid.arrange(
  grobs = c(depth_plots),
  nrow = 2,
  top = grid::textGrob(paste(species, "(", covs, ")"))
)
dev.off()
```

@@

```{r}
write_grads(signif(max(depth_model_list$adult$gradients), digits = 8), paste0(spp, "-mat"))
write_grads(signif(max(depth_model_list$imm$gradients), digits = 8), paste0(spp, "-imm"))
write_grads(signif(max(depth_model_list$total$gradients), digits = 8), paste0(spp, "-total"))
```

@@

```{r eval = FALSE}
try({
  if(max(depth_model_list$imm$gradients)>0.001) {
    imm_biomass <- sdmTMB::run_extra_optimization(depth_model_list$imm, nlminb_loops = 2, newton_steps = 2)
    write_grads(signif(max(imm_biomass$gradients), digits = 8), paste0(spp, "-imm2"))
  }
})
```

```{r eval = FALSE}
try({
  if(max(depth_model_list$total$gradients)>0.001) {
    total_biomass <- sdmTMB::run_extra_optimization(depth_model_list$total, nlminb_loops = 2, newton_steps = 2)
    write_grads(signif(max(total_biomass$gradients), digits = 8), paste0(spp, "-total2"))
  }
})
```

@@

    saveRDS(imm_biomass,
      file = paste0("data/", spp, "/mod-imm-biomass-", 
        spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

  } else {
    saveRDS(total_biomass, file = paste0(
      "data/", spp,
      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))
  }
```

>>>>

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    # species: "Redbanded Rockfish"
    #species: "Pacific Ocean Perch"
    #species: "Pacific Cod"
    #species: "Petrale Sole"
    species: "Canary Rockfish"
    # # species: "Redstripe Rockfish"
    # species: "Shortbelly Rockfish"
    # species: "Lingcod"
    region: "All synoptic surveys"
    knots: 500
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    #covs: "-ssid-only" # covariate string used to name model rds 
    covs: "-tv-depth-only"
    #covs: "-trawled-ssid"
    update_model: TRUE
    update_predictions: FALSE
    update_model_check: FALSE
---

@@

covars <- readRDS("data/event-covariates.rds") %>% select(-geometry)
data <- dplyr::left_join(biomass, covars)

@@

if (region == "All synoptic surveys") {
  # spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
  spde <- sdmTMB::make_mesh(data, c("X", "Y"), n_knots = knots, type = "kmeans")
}

@@

    saveRDS(adult_biomass,
      file = paste0(
        "data/", spp,
        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-new.rds"
      )
    )

   # try({ imm_formula <- as.formula(paste(
   #    "imm_density ~ 0 + as.factor(year)", covariates, ""
   #  ))
   # 
   #  imm_biomass <- sdmTMB::sdmTMB(
   #    data = data,
   #    imm_formula,
   #    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
   #    time = "year", spde = spde,
   #    family = tweedie(link = "log"),
   #    # ar1_fields = TRUE, # all off for final version
   #    # reml = TRUE,
   #    include_spatial = TRUE,
   #    nlminb_loops = 2, 
   #    newton_steps = 2,
   #    enable_priors = priors,
   #    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
   #    silent = FALSE
   #  )
   # 
   #  saveRDS(imm_biomass,
   #    file = paste0("data/", spp, "/mod-imm-biomass-", 
   #      spp, covs, "-", ssid_string, "-new.rds"))
   # })

  } else {
    # dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
    # 
    # total_biomass <- sdmTMB::sdmTMB(
    #   data = data,
    #   dens_formula,
    #   time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
    #   time = "year", spde = spde,
    #   family = tweedie(link = "log"),
    #   # ar1_fields = TRUE, # all off for final version
    #   include_spatial = TRUE,
    #   nlminb_loops = 2, 
    #   newton_steps = 2,
    #   enable_priors = priors,
    #   control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    #   silent = FALSE
    # )
    # 
    # saveRDS(total_biomass, file = paste0(
    #   "data/", spp,
    #   "/model-total-biomass-", spp, covs, "-", ssid_string, "-new.rds"
    # ))
  }
  tictoc::toc()
}
```

@@

  try({
    imm_biomass <- readRDS(paste0(
      "data/", spp,
      "/mod-imm-biomass-", spp, covs, "-", ssid_string, ".rds"
    ))

@@

    imm_predictions <- predict(imm_biomass, newdata = nd)
    saveRDS(imm_predictions, file = paste0(
      "data/", spp,
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass.rds"
    ))
  })
} else {
  total_biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
  ))

@@

if(params$update_model_check) {
try({
  biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, ".rds"
  ))
})

try({
  biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
  ))
})

@@

saveRDS(point_predictions, file = paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, ".rds"
))
}
```

```{r}
depth_only_predictions <- readRDS(paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, ".rds"
))

@@

try({
  imm_biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-1n3n4n16.rds"
  ))
  # imm_biomass<-sdmTMB:::update_model(imm_biomass)
  # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
})

if (!exists("adult_biomass")) {
  try({
    total_biomass <- readRDS(paste0(
      "data/", spp,
      "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
    ))
  })
}

@@

```{r}
png(
  file = paste0(
    "figs/", spp,
    "/depth-", spp, covs, "-no-AR1.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 6
)
gridExtra::grid.arrange(
  grobs = c(depth_plots),
  nrow = 2,
  top = grid::textGrob(paste(species, "(", covs, ")"))
)
dev.off()
```

@@

```{r}
write_grads(signif(max(depth_model_list$adult$gradients), digits = 8), paste0(spp, "-mat"))
write_grads(signif(max(depth_model_list$imm$gradients), digits = 8), paste0(spp, "-imm-redo"))
write_grads(signif(max(depth_model_list$total$gradients), digits = 8), paste0(spp, "-total-redo"))
```

@@

```{r eval = FALSE}
try({
  if(max(depth_model_list$imm$gradients)>0.001) {
    imm_biomass <- sdmTMB::run_extra_optimization(depth_model_list$imm, nlminb_loops = 2, newton_steps = 2)
    write_grads(signif(max(imm_biomass$gradients), digits = 8), paste0(spp, "-imm3"))
  }
})
```

```{r eval = FALSE}
try({
  if(max(depth_model_list$total$gradients)>0.001) {
    total_biomass <- sdmTMB::run_extra_optimization(depth_model_list$total, nlminb_loops = 2, newton_steps = 2)
    write_grads(signif(max(total_biomass$gradients), digits = 8), paste0(spp, "-total3"))
  }
})
```

@@

    saveRDS(imm_biomass,
      file = paste0("data/", spp, "/mod-imm-biomass-", 
        spp, covs, "-", ssid_string, ".rds"))

  } else {
    saveRDS(total_biomass, file = paste0(
      "data/", spp,
      "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
    ))
  }
```
